---
title: "Choropleth Mapping"
subtitle: "Geospatial Data"
editor: visual
date: 4 Dec 2023
date-modified: "`r Sys.Date()`"
code-copy: true
execute: 
  echo: true
  eval: true
  warning: false
  error: false
website: 
    google-analytics: "G-SGM5SNE7BQ"
format:
  html:
    code-overflow: wrap
    code-fold: false
    code-summary: "Show the code"
    css: styles.css
---

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**CREDIT**]{style="color:#3a9c88"}

Once again, the credit for the code chunks and the majority content of this post to [Dr. Kam Tin Seong](https://faculty.smu.edu.sg/profile/kam-tin-seong-486 "Geospatial Analytics Lecturer"), my geospatial analytics lecturer at the School of Computing and Information Systems, Singapore Management University. He has helped me tremendously, from having zero knowledge and programming background to comfortably using R for geospatial analysis.
:::

This post explores the functions in R to plot choropleth maps.

<br>

# **1. INITIAL**

## **1.1 Loading R Packages**

The following are the packages required for the hands-on exercise :

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[***p_load( )*** - pacman -]{style="color:#d46e15"} to load packages into R environment. This function will attempt to install the package from CRAN or the pacman repository list if it is not installed.
:::

```{r}
pacman::p_load(sf, tmap, tidyverse)
```

<br>

## **1.2 Data Importation**

### 1.2.1 Import Geospatial Data

Import **Singapore [Urban Redevelopment Authority](https://www.ura.gov.sg/Corporate/Media-Room/Media-Releases/pr14-33)(URA) Master Plan 2014 Subzone Boundary** from [data.gov.sg](#0) into **`mpsz`** simple feature data frame -

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[***st_read( )*** - sf-]{style="color:#d46e15"} to read the shapefile into simple features data frame format with 2 arguments below :

-   **dsn** - define data path.

-   **layer** - provide shapefile name.
:::

```{r}
mpsz = st_read(dsn = "data/geospatial",                
               layer = "MP14_SUBZONE_WEB_PL")
```

### 1.2.2 Import Aspatial Data

Import **Singapore [Urban Redevelopment Authority](https://www.ura.gov.sg/Corporate/Media-Room/Media-Releases/pr14-33)(URA) Master Plan 2014 Subzone Boundary** from [data.gov.sg](#0) into **`mpsz`** simple feature data frame -

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[***read_csv( )*** - readr -]{style="color:#d46e15"} to read a CSV file.
:::

```{r}
popdata <- read_csv("data/aspatial/respopagesextod2011to2020.csv")
```

<br>

## **1.3 Data Inspection**

### 1.3.1 Retrieve Information

```{r}
glimpse(popdata,60)
```

<br>

# **2. DATA PROCESSING**

## **2.1 Data Wrangling**

Following are the additional variables will be needed for later choropleth plotting -

-   YOUNG : age group 0 to 4 until age group 20 to 24.

-   ECONOMY ACTIVE : age group 25 - 29 until age group 60 - 64.

-   AGED : age group 65 and above.

-   TOTAL : all age groups.

-   DEPENDENCY : the ratio between young & aged against the economy active group.

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[[***pivot_wider( )***](https://tidyr.tidyverse.org/reference/pivot_wider.html) - tidyr -]{style="color:#d46e15"} to create new columns with column name based on `AG` & values based on `pop`.

[***mutate( )*** - dplyr -]{style="color:#d46e15"} to create new columns as mentioned above.

[***filter( )*** - dplyr -]{style="color:#d46e15"} to keep only `2020` for `Time` attribute.

[***group\_ by( )*** - dplyr -]{style="color:#d46e15"} to group 3 variables - PA, SZ and AG.

[***select( )*** - dplyr -]{style="color:#d46e15"} to keep only variables that mentioned.
:::

```{r}
popdata2020 <- popdata %>%
  filter(Time == 2020) %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from = AG, values_from = POP) %>%
  mutate(YOUNG = rowSums(.[3:6]) + rowSums(.[12])) %>%
  mutate(`ECONOMY ACTIVE` = rowSums(.[7:11]) + rowSums(.[13:15])) %>% 
  mutate(`AGED` = rowSums(.[16:21])) %>%
  mutate(`TOTAL` = rowSums(.[3:21])) %>%
  mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)/`ECONOMY ACTIVE`) %>%
  select(`PA`,`SZ`,`YOUNG`,`ECONOMY ACTIVE`,`AGED`,`TOTAL`,`DEPENDENCY`)
```

<br>

## 2.2 Merging Geospatial & Aspatial Data

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[[***mutate_at( )***](https://dplyr.tidyverse.org/reference/mutate_all.html?q=mutate_at#grouping-variables) - dplyr -]{style="color:#d46e15"} to modify specified columns `PA` and `SZ` in *`popdata2020`*.

-   `.fun` argument to indicates the function, i.e. `touper( )` applying on the specified columns. This function is to convert all text to uppercase in order to match the uppercase in the geospatial dataset.

[[***left_join( )***](https://dplyr.tidyverse.org/reference/mutate-joins.html) - dplyr -]{style="color:#d46e15"} to keep all observations in `mpsz`. This is an `outer joins`.

[***write_rds( )*** - readr -]{style="color:#d46e15"} to serialise `mpsz_pop2020`. By default, no compression as space is cheaper than time.
:::

```{r}
popdata2020 <- popdata2020 %>%
  mutate_at(.vars = vars(PA, SZ),
            .funs = list(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)
```

```{r}
mpsz_pop2020 <- left_join(mpsz, popdata2020,
                          by = c("SUBZONE_N" = "SZ"))
```

```{r}
write_rds(mpsz_pop2020, "data/mpszpop2020.rds")
```

<br>

# 3. CHOROPLETH MAPPING

## 3.1 Mapping with tmap

### 3.1.1 Plot with `qtm( )`

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[[***qtm( )***](https://www.rdocumentation.org/packages/tmap/versions/3.3-4/topics/qtm) - tmap -]{style="color:#d46e15"} to plot a thematic map quickly. However, the aesthetic layers are

-   fill - to specify which variable to be used for the choropleth. Set it `NULL` if just to draw the polygon borders.
:::

```{r}
#| fig-width: 8
#| fig-asp: 0.68
qtm(mpsz_pop2020,
    fill = "DEPENDENCY")
```

```{r}
#| fig-width: 8
#| fig-asp: 0.68
qtm(mpsz_pop2020,
    fill = NULL)
```

### 3.1.1 Plot Cartographic Furniture with `tmap` elements

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[[***tm_shape( )***](https://www.rdocumentation.org/packages/tmap/versions/3.3-4/topics/tm_shape) - tmap -]{style="color:#d46e15"} to specify shape object, which in this case is *`mpsz_pop2020`*. This is to set the bounding box for the plot.

[[***tm_fill( )***](https://www.rdocumentation.org/packages/tmap/versions/3.3-4/topics/tm_fill) - tmap -]{style="color:#d46e15"} to specify which column's value to be used for filling the polygons.

-   plotting options for `style` argument :

    -   Each unique value will be a distinct category - **`cat`**

    -   [Discrete gradient](https://r-spatial.github.io/classInt/reference/classIntervals.html) options - **`fixed`**, **`sd`**, **`equal`**, **`pretty`**, **`quantile`**, **`kmeans`**, **`hclust`**, **`bclust`**, **`fisher`**, **`jenks`**, **`dpih`**, **`headtails`**

    -   Continuous gradient options - `cont`, `order` & `log10`

    -   Log scale - **`log10_pretty`**

-   When the style is `fixed`, `breaks` argument should be specified.

-   colour options for `palette` argument : if not sure which palette to use, can input this code chunk at the console for references - `tmaptools::palette_explorer()`

[***tm_layout( )*** - tmap -]{style="color:#d46e15"} to specify title, margins, aspect ratio, frame, legend, etc..

-   `frame` by default is TRUE.

-   `frame.lwd` specify the width of the frame.

-   `asp` - if the aspect ratio is set to '0', the ratio will be adjusted to the aspect ratio of the device.

[***tm_borders( )*** - tmap -]{style="color:#d46e15"} to define the line colour, width, type for polygons' borders.

-   `col` to indicate line colour.

-   `lwd` to specify the line width.

-   `lty` to specify the line type such as `blank`, `dashed`, `dotted`, `dotdash`, `longdash`, `twodash`, `solid`

-   `alpha` to indicate the transparency level. Value between 0 (transparent) and 1 (not transparent). By default is 1.

[[***tm_compass( )***](https://www.rdocumentation.org/packages/tmap/versions/3.3-4/topics/tm_compass) - tmap -]{style="color:#d46e15"} to create compass.

-   Options available for `type` are `arrow`, `4star`, `8star`, `radar`, `rose`. By default is arrow.

[[***tm_scale_bar( )***](https://www.rdocumentation.org/packages/tmap/versions/3.3-4/topics/tm_scale_bar) - tmap -]{style="color:#d46e15"} to create scale bar.

-   `color.dark` - by default is set to black.

-   `color.light` - by default is set to white.

[[***tm_grid( )***](https://www.rdocumentation.org/packages/tmap/versions/3.3-4/topics/tm_grid) - tmap -]{style="color:#d46e15"} to draw graticule lines.
:::

::: panel-tabset
## Style : quantile

NOTE : The default colour scheme is `YlOrRd`, missing value will be in grey and default interval binning based on `pretty`.

```{r}
tmap_mode("plot")
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame.lwd = 5,
            asp = 0) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```

border colour to `white`, number of classes, n to `10`

```{r}
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          n = 10,
          style = "quantile", 
          palette = "Blues",
          title = "Dependency ratio") +
  tm_layout(main.title = "Test Elements",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame.lwd = 5,
            asp = 0) +
  tm_borders(alpha = 0.5, col = 'white') +
  tm_compass(type = "8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```

## sd

Apart from style changed to `sd`, palette to `OrRd`, borders to `dashed`, compass to `arrow`, and additional arguments for `scale_bar`.

```{r}
#| fig-width: 8
#| fig-asp: 0.68
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "sd", 
          palette = "OrRd",
          title = "Dependency ratio") +
  tm_layout(main.title = "Test Elements",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame.lwd = 5,
            asp = 0) +
  tm_borders(alpha = 0.5, lty = 'dashed') +
  tm_compass(type = "arrow", size = 2) +
  tm_scale_bar(color.dark = 'red', color.light = 'orange') +
  tm_grid(alpha = 0.2)
```

## equal

Style to `equal`, palette to `Greys`, borders to `dotted` and compass to `radar` .

NOTE : be careful between `Greys` and `grey` - the output is entirely different!!!

```{r}
#| fig-width: 8
#| fig-asp: 0.68
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "equal", 
          palette = "Greys",
          title = "Dependency ratio") +
  tm_layout(main.title = "Test Elements",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame.lwd = 5,
            asp = 0) +
  tm_borders(alpha = 0.5, lty = 'dotted') +
  tm_compass(type = "radar", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```

```{r}
#| fig-width: 8
#| fig-asp: 0.68
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "equal", 
          palette = "grey",
          title = "Dependency ratio") +
  tm_layout(main.title = "Test Elements",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame.lwd = 5,
            asp = 0) +
  tm_borders(alpha = 0.5, lty = 'dotted') +
  tm_compass(type = "radar", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```

## kmeans

Style to `kmeans` , palette to `YlOrRd`, borders to `longdash` and compass to `rose`.

```{r}
#| fig-width: 8
#| fig-asp: 0.68
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "kmeans",
          palette = "YlOrRd",
          title = "Dependency ratio") +
  tm_layout(main.title = "Test Elements",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame.lwd = 5,
            asp = 0) +
  tm_borders(alpha = 0.5, lty = 'longdash') +
  tm_compass(type = "rose", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```

## hclust

Style to `hclust` and border to `blank`, and test default without `palette` argument.

```{r}
#| fig-width: 8
#| fig-asp: 0.68
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "hclust",
          title = "Dependency ratio") +
  tm_layout(main.title = "Test Elements",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame.lwd = 5,
            asp = 0) +
  tm_borders(alpha = 0.5, lty = 'blank') +
  tm_compass(type = "rose", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```

## fisher

Style to `fisher`. The output same as `jenks`

```{r}
#| fig-width: 8
#| fig-asp: 0.68
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "fisher",
          title = "Dependency ratio") +
  tm_layout(main.title = "Test Elements",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame.lwd = 5,
            asp = 0) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "rose", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```

## headtails

Style to `headtails`.

```{r}
#| fig-width: 8
#| fig-asp: 0.68
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "headtails",
          title = "Dependency ratio") +
  tm_layout(main.title = "Test Elements",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame.lwd = 5,
            asp = 0) +
  tm_borders(alpha = 0.5) +
  tm_compass(type = "rose", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha = 0.2)
```
:::

### 3.1.2 Binning Customisation

```{r}
summary(mpsz_pop2020$DEPENDENCY)
```

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          breaks = c(0, 0.60, 0.70, 0.80, 0.90, 19.0)) +
  tm_borders(alpha = 0.5)
```

### 3.1.3 Colour Reversion

When the context requires to highlight smaller values, add a '-' prefix.

::: panel-tabset
## Highlight High Values

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Greens") +
  tm_borders(alpha = 0.5)
```

## Low Values

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "-Greens") +
  tm_borders(alpha = 0.5)
```
:::

### 3.1.4 Additional Elements Map Layout

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "jenks", 
          palette = "Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 0.1) +
  tm_layout(main.title = "Distribution of Dependency Ratio by planning subzone \n(Jenks classification)",
            main.title.position = "center",
            main.title.size = 1,
            legend.height = 0.45, 
            legend.width = 0.35,
            legend.outside = FALSE,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5)
```

### 3.1.5 Map Style

Note that element is not referring to the `style` argument under `tm_fill`.

::: panel-tabset
## Classic

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("classic")
```

## Gray

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("gray")
```

## Natural

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("natural")
```

## Cobalt

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("cobalt")
```

## Col_blind

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("col_blind")
```

## Albatross

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("albatross")
```

## Beaver

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("beaver")
```

## BW

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("bw")
```

## Watercolor

```{r}
tm_shape(mpsz_pop2020)+
  tm_fill("DEPENDENCY", 
          style = "quantile", 
          palette = "-Greens") +
  tm_borders(alpha = 0.5) +
  tmap_style("watercolor")
```
:::

To reset the default style -

```{r}
tmap_style("white")
```

### 3.2.1 Facet Choropleth Maps

There are 3 ways to plot -

-   multiple values to at least one of the asthetic arguments;

-   define a group-by variable in `tm_facets( )`;

-   create multiple stand-alone maps with `tmap_arrange( )`.

#### 3.2.1.1 Multiple values in `tm_fill( )`

::: panel-tabset
## Palette : Same Colour

```{r}
#| fig-width: 8
#| fig-asp: 0.68
tm_shape(mpsz_pop2020)+
  tm_fill(c("YOUNG", "AGED"),
          style = "equal", 
          palette = "Blues") +
  tm_layout(legend.position = c("right", "bottom")) +
  tm_borders(alpha = 0.5)
```

## Different Colours

```{r}
#| fig-width: 8
#| fig-asp: 0.68
tm_shape(mpsz_pop2020)+ 
  tm_polygons(c("DEPENDENCY","AGED"),
          style = c("equal", "quantile"), 
          palette = list("Blues","Greens")) +
  tm_layout(legend.position = c("right", "bottom"))
```
:::

#### 3.2.1.2 With`tm_facets( )`

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[***tm_facets( )*** - tmap -]{style="color:#d46e15"} to plot multiple small choropleth maps.

-   `free.coords` - works along with `by` argument. By default is `TRUE`, alphabetically?

-   `drop.units` - By default `TRUE`. When `FALSE`, they are plotted where mapped aesthetics are regarded as missing values.
:::

::: panel-tabset
## DROP.UNITS : TRUE

```{r}
#| fig-width: 12
#| fig-asp: 0.68
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by = "REGION_N", 
            free.coords = TRUE, 
            drop.units = TRUE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```

## FALSE

```{r}
#| fig-width: 12
#| fig-asp: 0.68
tm_shape(mpsz_pop2020) +
  tm_fill("DEPENDENCY",
          style = "quantile",
          palette = "Blues",
          thres.poly = 0) + 
  tm_facets(by = "REGION_N", 
            free.coords = TRUE, 
            drop.units = FALSE) +
  tm_layout(legend.show = FALSE,
            title.position = c("center", "center"), 
            title.size = 20) +
  tm_borders(alpha = 0.5)
```
:::

#### 3.2.1.3 With`tmap_arrange( )`

::: {.callout-tip collapse="true" appearance="simple" icon="false"}
## [**⇳ Usage of the code chunk below :**]{style="color:#3a9c88"}

[[***tmap_arrange( )***](https://www.rdocumentation.org/packages/tmap/versions/3.3-4/topics/tmap_arrange) - tmap -]{style="color:#d46e15"} to arrange multiple stand-alone maps in grid layout.

-   `sync` - navigation in view mode (zooming and panning) be synchronised. By default is `FALSE`.
:::

```{r}
#| fig-width: 14
#| fig-asp: 0.68
youngmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("YOUNG", 
              n = 4,
              style = "quantile", 
              palette = "Blues") +
  tm_layout(legend.height = 0.42, 
            legend.width = 0.38,
            frame.lwd = 3)

agedmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("AGED",
              n = 4,
              style = "quantile", 
              palette = "Blues") +
  tm_layout(legend.height = 0.42, 
            legend.width = 0.38,
            frame.lwd = 3)

econactmap <- tm_shape(mpsz_pop2020)+ 
  tm_polygons("ECONOMY ACTIVE",
              n = 4,
              style = "quantile", 
              palette = "Blues") +
  tm_layout(legend.height = 0.42, 
            legend.width = 0.38,
            frame.lwd = 3)

tmap_arrange(youngmap, agedmap, econactmap, 
             heights = 1, widths = 1, asp = 2, ncol = 2, nrow = 2, sync = TRUE)
```

### 3.2.2 Mapping specific Spatial Object for Specific Purposes

NOTE : When `tmap_mode` set to `view`, the histogram will not be able to display correctly. So this function cannot be used along with `legend.hist`.

::: panel-tabset
## NON-INTERACTIVE

```{r}
#| fig-width: 12
#| fig-asp: 0.68
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N == "CENTRAL REGION",]) +
  tm_fill("DEPENDENCY",
          style = "quantile", 
          palette = "-Blues", 
          legend.hist = TRUE, 
          legend.is.portrait = TRUE,
          legend.hist.z = 1) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.8, 
            legend.width = 5.0,
            legend.position = c("right", "bottom"),
            frame = FALSE) +
  tm_borders(alpha = 0.5, col = 'white')
```

## INTERACTIVE

```{r}
#| fig-width: 12
#| fig-asp: 0.68
tmap_mode("view")
tmap_options(check.and.fix = TRUE)
tm_shape(mpsz_pop2020[mpsz_pop2020$REGION_N == "CENTRAL REGION",]) +
  tm_fill("DEPENDENCY",
          style = "quantile", 
          palette = "-Blues") +
  tm_view(set.zoom.limits = c(12, 15)) +
  tm_layout(legend.outside = TRUE,
            legend.height = 0.6, 
            legend.width = 5.0) +
  tm_borders(alpha = 0.68, col = 'white')
```
:::
